Optimal (Multiway) Spatial Joins

### 2024-05-19 msj.tex
yufei

This paper revisits the {\em spatial join} (SJ) problem defined as follows. Let $k$ be a constant integer at least 2. In the {\em $k$-SJ} problem, the input comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$. The goal is to find all $k$-tuples $(r_1, r_2, ..., r_k)$ where
\myitems{
    \item $r_i \in R_i$ for each $i \in [1, k]$; and
    \item $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$, namely, the $k$ rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection.
}
We represent the set of these $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$, referred to as the {\em join result}. Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.

@@

SJ is a fundamental operation in spatial databases, which manage geometric entities such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc. It plays a crucial role in implementing the {\em filter-refinement mechanism}, the dominant approach in spatial databases for computing the overlays of different entities. Specifically, a geometric entity, at the lowest level, is usually modeled as a polygon. Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be very expensive when the polygons have complex boundaries. For better efficiency, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $P$, which is the smallest axis-parallel rectangle enclosing $P$. Two polygons $P_1$ and $P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ have a non-empty intersection. To overlay $k$ different sets of entities, filter-refinement first performs ``filtering'' by using an SJ to fetch all $k$-tuples $(r_1, r_2, ..., r_k)$ where each $r_i$ ($1 \le i \le k$) is the MBR of a polygon $P_i$ and a ``refines'' every such $k$-tuple by checking if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection.

@@

Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling the $k$-SJ problem directly, imply $k$-SJ algorithms with $O((n + \out) \cdot \polylog n)$ running time. Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates. Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ... \times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$. Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples. Tao and Yi \cite{ty22} presented a data structure of $O(n \polylog n)$ space achieving the purpose. Combining their structure with the planesweep technique, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.

>>>>

This paper revisits the {\em spatial join} (SJ) problem formulated as follows. Let $k \ge 2$ be a constant integer. In the {\em $k$-SJ} problem, the input comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$. The goal is to find all $k$-tuples $(r_1, r_2, ..., r_k)$ where
\myitems{
    \item $r_i \in R_i$ for each $i \in [1, k]$; and
    \item $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$, namely, the $k$ rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection.
}
We represent the set of $k$-tuples described above as $\J_k(R_1, R_2, ..., R_k)$, referred to as the {\em join result}. Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.

@@

SJ is a fundamental operation in spatial databases, which manage {\em geometric entities} such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc. It plays a crucial role in implementing the {\em filter-refinement mechanism}, the dominant approach for overlaying different entities in real systems. To understand why, first note that a geometric entity is typically modeled as a polygon. Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be exceedingly expensive when the polygons have complex boundaries. To mitigate the issue, a spatial database stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$. Let $G_1, G_2, ..., G_k$ be $k$ different sets of geometric entities, and $R_1, R_2, ..., R_k$ be the corresponding sets of MBRs. To compute the overlay information from $G_1, G_2, ..., G_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $G_i$ whose MBR is $r_i$.

%Two polygons $P_1$ and $P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ intersect.

@@

Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling $k$-SJ directly, imply $k$-SJ algorithms with $O((n + \out) \cdot \polylog n)$ running time. Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates. Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ... \times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$. Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples. Tao and Yi \cite{ty22} presented a data structure of $O(n \polylog n)$ space achieving the purpose. Combining their structure with the planesweep technique, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
